---
title: "Lab 1"
author: "Danielle Hoekstra"
date: "2023-01-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(sf)
library(tmap)
library(here)
```

### Read in the data
```{r}
sf_trees <- read.csv(here("data", "sf_trees", "sf_trees.csv"))
```

### plot the top 5 legal status
```{r}
top_5_status <- sf_trees %>% 
  group_by(legal_status) %>%  
  summarize(tree_count = n()) %>% 
  slice_max(tree_count, n = 5) %>%
  arrange(-tree_count)
  
tree_plot <- ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), 
                                y = tree_count)) + 
  geom_col() +
  labs(y = 'Tree count', x = 'Legal Status') +
  coord_flip() +
  theme_minimal()

tree_plot
```

### Keep obs where legal status is "Permitted Site" and caretaker is "MTA"
```{r}
permitted_mia <- sf_trees %>% 
  filter(legal_status == "Permitted Site" & caretaker == "MTA")
```

### Keep Blackwood acacia trees, keep only legal_status, date, latitude, longitude columns
```{r, warning=FALSE}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
  select(legal_status, date, latitude, longitude)

#str_detect for if the name like this is next to another scientific name
### now make a plot

ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude)) + geom_point()
```

### use tidyr:: separate() to separate one column into multiple (in this case, separate common and scientific name) and use tidyr :: unite() to bring them back together
```{r, warning=FALSE}
sf_tress_sep <- sf_trees %>% separate(species, into = c('spp_scientific', 'spp_common'), sep = " :: ")

sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:species, sep = '_NEW_')
```

### Make some actual maps!
```{r}
blackwood_acacia_sf <- blackwood_acacia %>% 
  drop_na(longitude, latitude) %>% 
  st_as_sf(coords = c('longitude', 'latitude'))

# now you have coordinates as points and can plot them using simple features tool

st_crs(blackwood_acacia_sf) <- 4326 #basically a coordinate reference system, basic latitude and longitude


```









